---
title: "502_Project1"
author: "Ming Cai & Lamia Benyamine"
date: "2024-10-06"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# INTRODUCTION

There are many methods to approximate the probability of success, p,
from a Binomial random sample, however, not all have the same coverage
probabilities or average length of the confidence interval. This report
will review several different methods using R and compare the
performance of the confidence intervals for various sample sizes and
probabilities for these three properties:

-   The proportion of intervals that capture the true value

-   The proportion that miss above and proportion that miss below

-   The average length of the interval

We will use data collected from 1,00 random samples from a binomial
distribution where n varies from 15, 13, to 100 and p varies from 0.01
to 0.99 across each approximation method. The methods reviewed in this
report are: Wald interval Adjusted Wald interval Clopper-Pearson (exact)
interval Score interval Raw percentile interval using a parametric
bootstrap Bootstrap t interval using a parametric bootstrap

# SIMULATION METHODS

## Wald Interval
The formula of calculating Wald interval is:

$$
\hat{p} - z_{1-\frac{\alpha}{2}}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}} < p < \hat{p} + z_{1-\frac{\alpha}{2}}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}},
$$

where the $z_c$ denotes the $c$ quantile of the standard normal distribution. 

To calculate the Wald interval for $p$, we need to use sample's probability $\hat{p}$, sample size $n$ and the confidence level $1-\alpha$. The function of calculating Wald Interval is given below:

```{r}
WaldCI <- function(y, n, alpha = 0.05){
  p_hat = y/n  # Estimate of the probability of success
  # Return lower and upper bound of Wald CI
  c("wald ci lower" = p_hat - qnorm(1-alpha/2)*sqrt((p_hat*(1-p_hat))/n),
    "wald ci upper" = p_hat + qnorm(1-alpha/2)*sqrt((p_hat*(1-p_hat))/n))
}
```
We use $\alpha = 0.05$ as default in all calculation.

## Adjusted Wald interval

## Clopper-Pearson (exact) interval

## Score interval

Score tests, and in particular their standard errors, are based on the
log-likelihood at the null hypothesis value of the parameter, whereas
Wald tests are based on the log-likelihood at the maximum likelihood
estimate.

```{r}
#example code block

```

## Raw percentile interval using a parametric bootstrap

This method uses the empirical 1 − α/2 quantile for the lower value and
α/2 quantile for the upper value as the confidence interval of the
bootstrap distribution. For a parametric bootstrap, we will resample the
data from a fitted curve, in this case, a binomial distribution.

## Bootstrap t interval using a parametric bootstrap

This method creates a ‘t-type’ statistic to use a pivot.

# RESULTS

# CONCLUSION
